---
title: "Day 6 Homework"
output: html_document
---

1. Think carefully about your data.  Think about something you would like to do multiple times, to subsets of your data.  

I would like the mean accuracy and reaction time for each context (mean accuracy for initial, medial, and final contexts)

2. Write out (in plain English, in the notebook) what you are trying to do.

First, I would subset the data according to the level of the 'context' factor. Then, I would calculate the mean of the 'meanRT' vector and the 'meanACC' vector.  After doing this three times, I would combine the three factors to have a dataframe with the mean reaction time and accuracy for each context. 

3. Write some "pseudo-code" comments in an R chunk, as an outline
```{r echo=FALSE}
for(this.counter 1:3){
  the groups are the levels of the 'context' factor
  rt.vector <- subset the data where the row is the context level and the column is rt
  acc.vector <- subset the data where the row is the context level and the column is accuracy
  mean.df [1, this.counter] <- mean of the rt.vector
  mean.df [2, this.counter] <- mean of the acc.vector
}

```


4. Write some (real) code that will do what you want ONCE, on one subset of the data.

```{r}
ax <- read.csv("ax_dprime_cloze_AL.csv")
summary(ax$context)
ax.final.acc <- ax[ax$context=="Final", "meanACC"]
ax.final.rt <- ax[ax$context=="Final", "meanRT"]
mean.df <- data.frame(final=vector(length=2), initial = vector(length=2), medial=vector(length=2))
mean.df[1,1] <- mean(ax.final.acc)
mean.df[2,1] <- mean(ax.final.rt)
mean.df

```


5. Copy that code chunk, and convert it into a loop, which will do what you want over all the groups, concatenating the results.

```{r}
mean.df.all <- data.frame(final=vector(length=2), initial = vector(length=2), medial=vector(length=2))

for(this.counter in 1:3){
 this.group <- c("Final", "Initial", "Medial")[this.counter]
 acc.vector <- ax[ax$context == this.group, "meanACC"]
 rt.vector<- ax[ax$context == this.group, "meanRT"]
 mean.df.all[1, this.counter] <- mean(acc.vector)
 mean.df.all[2, this.counter] <- mean(rt.vector)
}  
mean.df.all
this.counter

```


6. Copy that loop again, and at a condition (`if`) statement to add some further control about ignoring some groups or some other conditional operation.

Only go through the if loop for reaction times greater than 750ms

```{r}

if(x > 5) {
  x <- x^2
}

ax <- read.csv("ax_dprime_cloze_AL.csv")
# Attempt 1 - the if statement isn't applying
rm(mean.df.reduced)
mean.df.reduced <- data.frame(final=vector(length=2), initial = vector(length=2), medial=vector(length=2))
if(ax$meanRT < 750){
  for(this.counter in 1:3){
    this.group <- c("Final", "Initial", "Medial")[this.counter]
    acc.vector <- ax[ax$context == this.group, "meanACC"]
    rt.vector<- ax[ax$context == this.group, "meanRT"]
    mean.df.reduced[1, this.counter] <- mean(acc.vector, na.rm=TRUE)
    mean.df.reduced[2, this.counter] <- mean(rt.vector, na.rm=TRUE)
}}  
mean.df.reduced
mean.df.all
this.counter

# Attempt 2 - now all vectors are the last iteration of the function (i.e., the final and initial are the same as the medial means)
rm(mean.df.reduced)
mean.df.reduced <- data.frame(final=vector(length=2), initial = vector(length=2), medial=vector(length=2))

for(this.counter in 1:3){
  if(ax$meanRT[this.counter] > 750){
    this.group <- c("Final", "Initial", "Medial")[this.counter]
    acc.vector2 <- ax[ax$context == this.group, "meanACC"]
    rt.vector2<- ax[ax$context == this.group, "meanRT"]
    mean.df.reduced[1, this.counter] <- mean(acc.vector)
    mean.df.reduced[2, this.counter] <- mean(rt.vector)
}}  
mean.df.reduced
mean.df.all
this.counter
this.group

# Attempt 3 
rm(mean.df.reduced)
mean.df.reduced <- data.frame(final=vector(length=2), initial = vector(length=2), medial=vector(length=2))

for(this.counter in 1:3){
    this.group <- c("Final", "Initial", "Medial")[this.counter]
      if(ax$meanRT[this.counter] > 750){
    acc.vector2 <- ax[ax$context == this.group, "meanACC"]
    rt.vector2<- ax[ax$context == this.group, "meanRT"]
    mean.df.reduced[1, this.counter] <- mean(acc.vector)
    mean.df.reduced[2, this.counter] <- mean(rt.vector)
}}  
mean.df.reduced
mean.df.all
this.counter
this.group

# Attempt 4 - if statement didn't apply
rm(mean.df.reduced)
ax <- read.csv("ax_dprime_cloze_AL.csv")

if(any(ax$meanRT) < 750){
  ax$meanRT <- NA
  } else {
  ax$meanRT <- ax$meanRT
}

summary(ax$meanRT)
summary(ax)
for(this.counter in 1:3){
    this.group <- c("Final", "Initial", "Medial")[this.counter]
    acc.vector2 <- ax[ax$context == this.group, "meanACC"]
    rt.vector2<- ax[ax$context == this.group, "meanRT"]
    mean.df.reduced[1, this.counter] <- mean(acc.vector, na.rm=TRUE)
    mean.df.reduced[2, this.counter] <- mean(rt.vector, na.rm=TRUE)
}  

# Attempt 5 - this works but only uses an if statement
rm(mean.df.reduced)
ax <- read.csv("ax_dprime_cloze_AL.csv")

ax$meanRT <- ifelse(ax$meanRT< 750, NA, ax$meanRT)

mean.df.reduced <- data.frame(final=vector(length=2), initial = vector(length=2), medial=vector(length=2))


summary(ax$meanRT)
for(this.counter in 1:3){
 this.group <- c("Final", "Initial", "Medial")[this.counter]
 acc.vector <- ax[ax$context == this.group, "meanACC"]
 rt.vector<- ax[ax$context == this.group, "meanRT"]
 mean.df.reduced[1, this.counter] <- mean(acc.vector, na.rm=TRUE)
 mean.df.reduced[2, this.counter] <- mean(rt.vector, na.rm=TRUE)
} 

mean.df.reduced
mean.df.all
this.counter
```

