---
title: "Day 7 Homework"
author: "Alia Lancaster"
output: html_document
---

1. Work on making progress on your Final Project.
Plan of attack + code for what I've previously done

Read in data and packages
```{r}
ax <- read.csv("ax_AL.csv")

library(ggplot2)
library(plyr)
```


1) create a column that is a factor similar to 'phoneme_pair' but that also takes into account the same pairs
- in previous script, created a column for each phoneme pair

```{r}
ax$Ah <- ifelse(ax$phoneme_pair=="A-h" | ax$phoneme_pair=="A-A-" | ax$phoneme_pair == "hh","Ah",0)
ax$Ah <- as.factor(ax$Ah)
summary(ax$Ah)
# repeat 26 times 

```

2) Add 4 factors for hits, misses, false alarms, and correct responses
```{r}
ax$HITS <- ifelse(ax$correct_response=="0" & ax$ACC==1,1,0)
ax$FA <- ifelse(ax$correct_response=="0" & ax$ACC==0,1,0)
ax$m <- ifelse(ax$correct_response=="1" & ax$ACC==0,1,0)
ax$cr <- ifelse(ax$correct_response=="1" & ax$ACC==1,1,0)
summary(ax)
```


3) calculate number of hits and false alarms for each subject, pair, and context.

- accomplished by subestting by pair, then dply for each subject, context, and speaker
```{r}
#subset
ax.Ah <- ax[ax$Ah == "Ah", ]
ax.Ah <- droplevels(ax.Ah)

# ddply
dprime.table.Ah <- ddply (ax.Ah, c ("SubjectID", "context", "spkr"), summarize, meanACC = mean(ACC), meanRT = mean(RT), Hits = sum(HITS), FAlarms = sum(FA), miss = sum(m), cr= sum(cr), n = length(RT))

head(dprime.table.Ah)
summary(dprime.table.Ah)
count(dprime.table.Ah$FAlarms)
count(dprime.table.Ah$Hits)

# repeat all code in this chunk 26 times

```

4) make a table with the frequency of special cases
- Haven't done this before

```{r}
#pseudo-code
save counts of FA and hits
delete all cases except of 0 and 1
put into a data frame and record the pair

# real code
```


5) Dealt with special cases, calculate proportion of hits and false alarms, then d-prime.

```{r}
#replacing 0s and 1s 
attach(dprime.table.Ah)
dprime.table.Ah$pH <- ifelse((dprime.table.Ah$Hits/(dprime.table.Ah$Hits + dprime.table.Ah$miss)) ==1, 0.875, ifelse((dprime.table.Ah$Hits/(dprime.table.Ah$Hits + dprime.table.Ah$miss))==0, 0.125, dprime.table.Ah$Hits/(dprime.table.Ah$Hits + dprime.table.Ah$miss))) 
dprime.table.Ah$pFA <- ifelse((dprime.table.Ah$FAlarms/(dprime.table.Ah$FAlarms + dprime.table.Ah$cr)) ==1, 0.986, ifelse((dprime.table.Ah$FAlarms/(dprime.table.Ah$FAlarms + dprime.table.Ah$cr))==0, 0.0135, dprime.table.Ah$FAlarms/(dprime.table.Ah$FAlarms + dprime.table.Ah$cr))) 
detach(dprime.table.Ah)
#dprime
dprime.table.Ah$dprime <- qnorm(dprime.table.Ah$pH)-qnorm(dprime.table.Ah$pFA)

dprime.table.Ah$pair <- "Ah"
dprime.table.Ah$pair <- as.factor(dprime.table.Ah$pair)
summary(dprime.table.Ah)

```

5b) Because I subset, I then put everything back together with rbind(), and then recreated the 'pu' and 'vvl' factors.

```{r}
#dprime.all1<- rbind (dprime.table.Ah, dprime.table.AH) 
# repeat 25 times

##### create columns for p/u and v/vl #####
#attach(dprime.all)
#dprime.all$pu[pair =="AH"] <- "pharyngeal"
#dprime.all$pu[pair =="Rx"] <- "uvular"
#dprime.all$pu[pair !="AH" & pair !="Rx"] <- "control"
#detach(dprime.all)
#dprime.all$pu <- as.factor(dprime.all$pu)
#summary(dprime.all)

#attach(dprime.all)
#dprime.all$vvl[pair =="AR"] <- "voiced"
#dprime.all$vvl[pair =="Hx"] <- "voiceless"
#dprime.all$vvl[pair !="AR" & pair !="Hx"] <- "ccontrol"
#detach(dprime.all)
#dprime.all$vvl <- as.factor(dprime.all$vvl)
#summary(dprime.all)

```

6) Use boxplots to view distribution of d-prime across contexts and by speaker
```{r}
ax.dprime <- read.csv("ax_dprime_cloze_AL.csv")
summary(ax.dprime)
#get rid of NAs
ax.dprime <- ax.dprime[!is.na(ax.dprime$dprime),]
#reorder context
ax.dprime$context <- factor(ax.dprime$context, levels = c("Initial", "Medial", "Final"))


ggplot(ax.dprime, aes(x=context, y=dprime, fill = speaker)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Context") + ylab("D-Prime") + theme_bw()

ggplot(ax.dprime, aes(x=speaker, y=dprime, fill = pu)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Context") + ylab("D-Prime") + theme_bw()

ggplot(ax.dprime, aes(x=speaker, y=dprime, fill = pu)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Context") + ylab("D-Prime") + theme_bw() + facet_grid (.~context)

ggplot(ax.dprime, aes(x=speaker, y=dprime, fill = vvl)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Context") + ylab("D-Prime") + theme_bw()

ggplot(ax.dprime, aes(x=speaker, y=dprime, fill = vvl)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Context") + ylab("D-Prime") + theme_bw() + facet_grid (.~context)

```


2. Document some road blocks you are running into.

4. Send Scott special topics you're intersted in.

3. Email them to me before 7am tomorrow. and repo name