---
title: "Day 9 Homework"
author: "Alia Lancaster"
output: html_document
---
packages and functions
```{r}
library(dplyr)
library(reshape2)
```


1. Write a function to do a bunch of clean-up and formatting to your data.

Note: I'm using the dataset without dprime, "ax_AL.csv".  Both datasets I've used should be in my repo.


```{r}
ax <- read.csv("ax_AL.csv")
summary(ax)


# what I want my function to do

# create hits, FA, misses, and cr columns then calculate proportions of hits and false alarms.  Just a skelton for now, since the pairs aren't quite right.
ax$HITS <- ifelse(ax$correct_response=="0" & ax$ACC==1,1,0)
ax$FA <- ifelse(ax$correct_response=="0" & ax$ACC==0,1,0)
ax$m <- ifelse(ax$correct_response=="1" & ax$ACC==0,1,0)
ax$cr <- ifelse(ax$correct_response=="1" & ax$ACC==1,1,0)
summary(ax)

#cbind
one <- vector(length=nrow(ax))
ax<- cbind(ax, one)
names(ax)

# function
make.newcol <- function(data){
  HITS <- ifelse(data$correct_response=="0" & data$ACC==1,1,0)
  FA <- ifelse(data$correct_response=="0" & data$ACC==0,1,0)
  m <- ifelse(data$correct_response=="1" & data$ACC==0,1,0)
  cr <- ifelse(data$correct_response=="1" & data$ACC==1,1,0)
  data <- cbind(data, HITS, FA, m, cr) 
}

```


2. Make sure this function works on your data.

```{r}
ax <- read.csv("ax_AL.csv")
ax <- make.newcol(ax)
summary(ax)

```


3. Get some subset of your data, or find a similar-but-not-identical data set.  Change up the column order.  Re-run your function on this new data.  Does it work?

The function works on my data because I used column names instead of numbers.

```{r}
#subset
ax.subset <- read.csv("ax_AL.csv")
names(ax.subset)
ax.subset <- ax[,c(2:8, 12:17)]
names(ax.subset)

# reorder columns in dataframe
ax.subset <- ax.subset[c(3,4, 1, 2, 13, 12, 11, 10, 9, 8, 7, 6, 5)]
names(ax.subset)

# does the function work? Yes!
ax.subset <- make.newcol(ax.subset)
summary(ax.subset)
```


4. Fix your function to be able to work on the new data.

5. Write a function to do some kind of summary stat, and use the `dplyr` functions to run this function over groups of your data.

 I wrote a function to calculate the proportion of two vectors to generate the of hits (hits/hits+ misses) and proportion of false alarms (fa/fa+cr).  I then used the dplyr functions to apply this function over each Subject/context group and add this vector to the existing dataframe. These new columns (pH and pFA) will be used later to calculate d-prime (but after a few more changes to the data).
 
 
```{r}
prop <- function(vector1, vector2){
  numer <- sum(vector1)
  denom <- sum(vector1) + sum(vector2)
  output <- numer/denom
  output
}

prop(ax$HITS, ax$m)

ax <- ax %>%
  group_by(SubjectID, context) %>%
  mutate(pH = prop(HITS, m), pFA = prop(FA, cr))

summary(ax)  

```

