---
title: "final project"
author: "Alia Lancaster"
output: html_document
---
Read in data and packages
```{r}
ax <- read.csv("ax_AL.csv")

library(ggplot2)
library(dplyr)
library(reshape2)
```

Goal: Calculate d-prime per subject, context (initial, medial, final), and pair (28 pairs of phonemes). Record the number of times the proportion of hits or misses is 0 or 1, then replace these special cases with values that represent a similar pattern of response but do not cause "Inf" or "Nan". Create boxplots to examine the distribution of d-prime.


Step 1) Create a new column for each phoneme pair that takes into account the same and different pairs.  For instance, the factor "phoneme_pair" includes "Ah", "AA", and "hh".  Create a column called "Ah" that records if a trial is any of those three pairs.  A separate column is needed for each pair because the same pairs overlap.  For instance, the same pairs for "Ah" and "Ax" both include "AA".  These new columns will be used in subsetting later on. 

```{r}
# make all of the strings in 'phoneme_pair' equal to 2 characters.  Take out or replace "-"
summary(ax$phoneme_pair)
ax$phoneme_pair <- factor(gsub("A-", "A", ax$phoneme_pair))
ax$phoneme_pair <- factor(gsub("h-", "H", ax$phoneme_pair))
summary(ax$phoneme_pair) 

#split by nothing and rbind to ax
sst <- do.call(rbind, strsplit(as.character(ax$phoneme_pair), ""))
phoneme1 <- sst[,1]
phoneme2 <- sst[,2]
ax <- cbind(ax, phoneme1, phoneme2)
names(ax)

# Create a new column for each pair, collapsing across same and different trials
paircol <- vector(length=length(ax$phoneme_pair))
for(pair in levels(droplevels(ax[ax$correct_response==0, "phoneme_pair"]))){
    pair1 <- strsplit(as.character(pair), "")[[1]][1]
    pair2 <- strsplit(as.character(pair), "")[[1]][2]
    paircol <- factor(ifelse(ax$phoneme_pair==pair, pair, ifelse(ax$correct_response==1 &  ax$phoneme1==pair1, pair, ifelse(ax$correct_response==1 & ax$phoneme1 ==pair2, pair, 0))))
    ax<- cbind(ax, paircol)
    names(ax)[names(ax)=="paircol"] <- pair
  }

```

Step 2) Add 4 factors for hits, misses, false alarms, and correct responses. 


```{r}
make.newcol <- function(data, x, x2){
  hits <- ifelse(data[,x]=="0" & data[,x2]==1,1,0)
  fa <- ifelse(data[,x]=="0" & data[,x2]==0,1,0)
  m <- ifelse(data[,x]=="1" & data[,x2]==0,1,0)
  cr <- ifelse(data[,x]=="1" & data[,x2]==1,1,0)
  data <- cbind(data, hits, fa, m, cr) 
}
ax <- make.newcol(ax, "correct_response", "ACC")
summary(ax)
```


Step 3) Calculate number of proportion of hits and false alarms for each pair, subject, context. There are 28 pairs, 32 subjects, and three contexts, so there should be 896 rows per context

```{r}
# proportion function

prop <- function(vector1, vector2){
  numer <- sum(vector1)
  denom <- sum(vector1) + sum(vector2)
  output <- numer/denom
  output
}
# for loop
dprime.all <- data.frame()
this.subset <- data.frame()
for(column in colnames(ax[, c(20:47)])){
  this.subset <- ax[ax[,column] == column, c("Item", "SubjectID", "context", "correct_response","RT", "ACC", "spkr", "pu", "vvl", "list", "hits", "fa", "m", "cr") ]
  this.subset <- droplevels(this.subset)
  this.sum <- this.subset %>%
    group_by(SubjectID, context, spkr) %>%
    summarize(pH = prop(hits, m), pFA = prop(fa, cr), pair = column, meanRT = mean(RT), meanACC= mean(ACC))
  dprime.all <- rbind(dprime.all, this.sum)
}

summary(dprime.all)
length(dprime.all$context)

#the rbind made some factors into characters, so change them back to factors
dprime.all$pair <- as.factor(dprime.all$pair)
dprime.all$pair <- as.factor(dprime.all$pair)

## add the pu and vvl factors back
dprime.all$pu <- as.factor(ifelse(dprime.all$pair=="AH", "p", ifelse(dprime.all$pair=="Rx", "u", "c")))
summary(dprime.all)

dprime.all$vvl <- as.factor(ifelse(dprime.all$pair=="AR", "v", ifelse(dprime.all$pair=="Hx", "vl", "c")))
summary(dprime.all)
```


Step 4) Make tables of frequency of special cases. The if/else loop in the function exists because pFA does not have any instances of 1, which makes the number of columns different in "final.df". 

```{r}

freq.tables <- function(data, groupvar, numvar){
  groupvar.sum<- xtabs(~data[, groupvar] + data[,numvar], data=data)
  groupvar.sum <- as.data.frame(groupvar.sum)
  groupvar.sum <- groupvar.sum[groupvar.sum[,2]==0 | groupvar.sum[, 2]==1, ]
  groupvar.cast <- dcast(groupvar.sum, groupvar.sum[,1] ~ groupvar.sum[,2], value.var="Freq")
  total <- xtabs(~data[, groupvar], data=data)
  total <- as.data.frame(total)
  final.df <- cbind(groupvar.cast, total[, 2])
  if ((length(names(final.df))) > 3){
  final.df$Perc0 <- round((final.df[, "0"]/final.df[,4])*100, digits=3)
  final.df$Perc1 <- round((final.df[, "1"]/final.df[,4])*100, digits=3)
  final.df$TotalPerc <- final.df$Perc0 + final.df$Perc1
  output <- final.df
  } else
  final.df$Perc0 <- round((final.df[, "0"]/final.df[,3])*100, digits=3)
  output <- final.df 
}

pair.pH <-freq.tables(dprime.all, "pair", "pH")
subject.pH <- freq.tables(dprime.all, "SubjectID", "pH")
subject.pFA <- freq.tables(dprime.all, "SubjectID", "pFA")
pair.pFA <- freq.tables(dprime.all, "pair", "pFA")

pair.pH
subject.pH
subject.pFA
pair.pFA
```


Step 5) Deal with special cases, then add d-prime. 
Replacements:
- hits: 4 trials, so 4/4 is the most possible hits and 0/4 is the least possible hits. The next higherst number of hits is 3/4 = .75, and the next lowest number of hits is 1/4 = .25.  Replace 1 with the value halfway between 1 and .75, which is .875. Replace 0 with the value halfway between 0 and .25, which is 0.125
- false alarms: 32 trials.  The next highest number of false alarms is 31/33 = 0.96875, and the next lowest number of false alarms is 1/32 = 0.03125. Replace 1 with the value halfway between 1 and .96875, which is 0.984375.  Replace 0 with the value halfway between 0 and 0.03125, which is 0.015625.   


```{r}

# the max number of different trials a person heard of a certain pair in a certain context
maxh = 4

# the max number of same trials a person heard of a certain pair in a certain context
maxfa = 32

dprime <- function(data, pH, pFA){
  data[, pH] <- ifelse(data[, pH]==1, (1-(1-((maxh-1)/maxh))/2), ifelse(data[, pH]==0, ((1/maxh)/2), data[, pH]))
  data[, pFA] <- ifelse(data[, pFA]==1, (1-(1-((maxfa-1)/maxfa))/2), ifelse(data[, pFA]==0, ((1/maxfa)/2), data[, pFA]))
  output <- qnorm(data[,pH]) - qnorm(data[,pFA])
  output
}


dprime.all$dprime <- dprime(dprime.all, "pH", "pFA")
summary(dprime.all)


```



Step 6) Use boxplots to view distribution of d-prime across contexts and by speaker

```{r}

#get rid of NAs for boxplots
dprime.all.noNA <- dprime.all[!is.na(dprime.all$dprime),]
#reorder context to initial, medial, final
dprime.all.noNA$context <- factor(dprime.all.noNA$context, levels = c("I", "M", "F"))


# dprime by speaker and context
ggplot(dprime.all.noNA, aes(x=context, y=dprime, fill = spkr)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Context") + ylab("D-Prime") + theme_bw()

# dprime by speaker and manner type (pharyngeal, uvular, control) 
ggplot(dprime.all.noNA, aes(x=spkr, y=dprime, fill = pu)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Speaker") + ylab("D-Prime") + theme_bw()

# dprime by speaker, context, and manner type (pharyngeal, uvular, control)
ggplot(dprime.all.noNA, aes(x=spkr, y=dprime, fill = pu)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Speaker") + ylab("D-Prime") + theme_bw() + facet_grid (.~context)

# dprime by speaker and voicing type (voiced, voiceless, control)
ggplot(dprime.all.noNA, aes(x=spkr, y=dprime, fill = vvl)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Speaker") + ylab("D-Prime") + theme_bw()

#dprime by speaker, context, and voicing type (voiced, voiceless, control)
ggplot(dprime.all.noNA, aes(x=spkr, y=dprime, fill = vvl)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Speaker") + ylab("D-Prime") + theme_bw() + facet_grid (.~context)

```

