---
title: "final project draft"
author: "Alia Lancaster"
output: html_document
---
Read in data and packages
```{r}
ax <- read.csv("ax_AL.csv")

library(ggplot2)
library(dplyr)
library(reshape2)
```


1) create a column that is a factor similar to 'phoneme_pair' but that also takes into account the same pairs
- in previous script, created a column for each phoneme pair
- plan: split 'phoneme_pair' column.  Then create a new column for for each pair, including the same pairs (need to because the same pair the the same for some pairs, e.g., Ah and Ax)

```{r}
length(ax$phoneme_pair) #21264
# make all of the strings in 'phoneme_pair' equal to 2 characters.  Take out or replace "-"
summary(ax$phoneme_pair)
ax$phoneme_pair <- factor(gsub("A-", "A", ax$phoneme_pair))
ax$phoneme_pair <- factor(gsub("h-", "H", ax$phoneme_pair))
summary(ax$phoneme_pair) # it worked!

#split by nothing and rbind to ax
sst <- do.call(rbind, strsplit(as.character(ax$phoneme_pair), ""))
length(sst[,2]) #21264
head(sst)
tail(sst)
phoneme1 <- sst[,1]
phoneme2 <- sst[,2]
ax <- cbind(ax, phoneme1, phoneme2)
names(ax)

# for every phoneme pair that is different (i.e., when correct_response is 0)
paircol <- vector(length=length(ax$phoneme_pair))
for(pair in levels(droplevels(ax[ax$correct_response==0, "phoneme_pair"]))){
    pair1 <- strsplit(as.character(pair), "")[[1]][1]
    pair2 <- strsplit(as.character(pair), "")[[1]][2]
    paircol <- factor(ifelse(ax$phoneme_pair==pair, pair, ifelse(ax$correct_response==1 &  ax$phoneme1==pair1, pair, ifelse(ax$correct_response==1 & ax$phoneme1 ==pair2, pair, 0))))
    ax<- cbind(ax, paircol)
    names(ax)[names(ax)=="paircol"] <- pair
  }

# check to see if it worked
names(ax)
length(ax$Ah)
summary(ax$Ah)
summary(ax$phoneme_pair) # from this, I see that there are 384 rows with Ah, 1341 for AA, and 1341 for hh.  This means that Ah should have 3069 rows (384+1341+1341) in ax$Ah 

```

2) Add 4 factors for hits, misses, false alarms, and correct responses
- Function for this from a homework (but improved to make more general by adding column names to function arguments).
- TO DO: add more to clean up data in any way?

```{r}
make.newcol <- function(data, x, x2){
  hits <- ifelse(data[,x]=="0" & data[,x2]==1,1,0)
  fa <- ifelse(data[,x]=="0" & data[,x2]==0,1,0)
  m <- ifelse(data[,x]=="1" & data[,x2]==0,1,0)
  cr <- ifelse(data[,x]=="1" & data[,x2]==1,1,0)
  data <- cbind(data, hits, fa, m, cr) 
}
ax <- make.newcol(ax, "correct_response", "ACC")
summary(ax)
```


3) calculate number of proportion of hits and false alarms for each pair, subject, context.

- accomplished by subestting by pair, then dply for each subject, context, and speaker
- can't use dplyr because Ah column overlaps with the AR column in same pairs, so have to do a for loop
- use the prop function I created for a homework to skip the step of calc the sum of each hit, fa, cr, m.  

To Do:
- there are no 1's in pFA for pair and subject. This messes up the number of columns after the cast.
- the length of the columns after the for loop is 4x longer (85249)

28 pairs, 32 subjects =896 per context

Proportion:
```{r}
dprime.all <- data.frame()
this.subset <- data.frame()

# for loop

for(column in colnames(ax[, c(20:47)])){
  this.subset <- ax[ax[,column] == column, c("Item", "SubjectID", "context", "correct_response","RT", "ACC", "spkr", "pu", "vvl", "list", "hits", "fa", "m", "cr") ]
  this.subset <- droplevels(this.subset)
  this.sum <- this.subset %>%
    group_by(SubjectID, context, spkr) %>%
    summarize(pH = prop(hits, m), pFA = prop(fa, cr), pair = column, meanRT = mean(RT), meanACC= mean(ACC))
  dprime.all <- rbind(dprime.all, this.sum)
}

summary(dprime.all)
#the rbind made some of my factors into characters, so change them back to factors
dprime.all$pair <- as.factor(dprime.all$pair)
dprime.all$pair <- as.factor(dprime.all$pair)

## add the pu and vvl factors back (I'll want them for analysis and graphing)
dprime.all$pu <- as.factor(ifelse(dprime.all$pair=="AH", "p", ifelse(dprime.all$pair=="Rx", "u", "c")))
summary(dprime.all)

dprime.all$vvl <- as.factor(ifelse(dprime.all$pair=="AR", "v", ifelse(dprime.all$pair=="Hx", "vl", "c")))
summary(dprime.all)




prop <- function(vector1, vector2){
  numer <- sum(vector1)
  denom <- sum(vector1) + sum(vector2)
  output <- numer/denom
  output
}


```

make tables of number of special cases
```{r eval=FALSE}


#pseudo-code
save counts of FA and hits
delete all cases except of 0 and 1
put into a data frame and record the pair
xtabs(~pH +pair, data=dprime.all)
# real code
pH <- dprime.all %>%
  group_by(SubjectID, context) %>%
  xtabs(~ pair + pH, data=.)
# create a function.  Want this for ph, pfa, use subject or context instead of pair


data <- dprime.all
groupvar <- "pair"
numvar <- "pFA"

myfunc <- function(data, groupvar, numvar){
  groupvar.sum<- xtabs(~data[, groupvar] + data[,numvar], data=data)
  groupvar.sum <- as.data.frame(groupvar.sum)
  groupvar.sum <- groupvar.sum[groupvar.sum[,2]==0 | groupvar.sum[, 2]==1, ]
  groupvar.cast <- dcast(groupvar.sum, groupvar.sum[,1] ~ groupvar.sum[,2], value.var="Freq")
  #names(groupvar.cast)[names(groupvar.cast)=="0"] <- paste("numvar", "0", sep = "")
  #names(groupvar.cast)[names(groupvar.cast)=="1"] <- paste("numvar", "1", sep = "")
  total <- xtabs(~data[, groupvar], data=data)
  total <- as.data.frame(total)
  final.df <- cbind(groupvar.cast, total[, 2])
  #final.df <- merge(groupvar.cast, total, by.y = groupvar.cast[, 1], by.x=total[, 1])
  #names(final.df)[names(final.df)=="Freq"] <- "total"
  if ((length(names(final.df))) > 3){
  final.df$Perc0 <- round((final.df[, "0"]/final.df[,4])*100, digits=3)
  final.df$Perc1 <- round((final.df[, "1"]/final.df[,4])*100, digits=3)
  final.df$TotalPerc <- final.df$Perc0 + final.df$Perc1
  output <- final.df
  } else
  final.df$Perc0 <- round((final.df[, "0"]/final.df[,3])*100, digits=3)
  output <- final.df 
}

pH2 <- myfunc(dprime.all, "pair", "pH")

#check to make sure that the function gets the same output as when I did it once:
pH<- xtabs(~pair + pH, data=dprime.all)
pH <- as.data.frame(pH)
pH <- pH[pH$pH==0 | pH$pH==1, ]
head(pH)
pH <- dcast(pH[1:56, ], pair ~ pH, value.var="Freq")
names(pH)[names(pH)=="0"] <- "pH0"
names(pH)[names(pH)=="1"] <- "pH1"
total <- xtabs(~pair, data=dprime.all)
total <- as.data.frame(total)
pH.total <- merge(pH, total)
names(pH.total)[names(pH.total)=="Freq"] <- "total"
pH.total$pH0Perc <- round((pH.total$pH0/pH.total$total)*100, digits=3)
pH.total$pH1Perc <- round((pH.total$pH1/pH.total$total)*100, digits=3)
pH.total$pHTotalPerc <- pH.total$pH0Perc + pH.total$pH1Perc

#these should have the same numbers
head(pH.total)
head(pH2)

#now run the function with pair/pFA, subjectID/pH, and subjectID/pFA:
pair.pH <-myfunc(dprime.all, "pair", "pH")
subject.pH <- myfunc(dprime.all, "SubjectID", "pH")
subject.pFA <- myfunc(dprime.all, "SubjectID", "pFA")
pair.pFA <- myfunc(dprime.all, "pair", "pFA")


```


5) Dealt with special cases, then d-prime.
Replacements:
- hits: 4 trials, so 4/4 is the most possible hits and 0/4 is the least possible hits. The next highers number of hits is 3/4 = .75, and the next lowest number of hits is 1/4 = .25.  Replace 1 with the value halfway between 1 and .75, which is .875. Replace 0 with the value halfway between 0 and .25, which is 0.125
- false alarms: 32 trials.  The next highest number of false alarms is 31/33 = 0.96875, and the next lowest number of false alarms is 1/32 = 0.03125. Replace 1 with the value halfway between 1 and .96875, which is 0.984375.  Replace 0 with the value halfway between 0 and 0.03125, which is 0.015625.   

- I know these numbers from the design. I made the number of trials into objects and put the calculations into the function.



```{r}

#function to replace special cases and calculate dprime:
# the max number of different trials a person heard of a certain pair in a certain context
maxh = 4

# the max number of same trials a person heard of a certain pair in a certain context
maxfa = 32

dprime <- function(data, pH, pFA){
  data[, pH] <- ifelse(data[, pH]==1, (1-(1-((maxh-1)/maxh))/2), ifelse(data[, pH]==0, ((1/maxh)/2), data[, pH]))
  data[, pFA] <- ifelse(data[, pFA]==1, (1-(1-((maxfa-1)/maxfa))/2), ifelse(data[, pFA]==0, ((1/maxfa)/2), data[, pFA]))
  output <- qnorm(data[,pH]) - qnorm(data[,pFA])
  output
}


#checks to make sure the function is working

pH1 <- c(.89, .69, .99, 1)
pH <- c(.89, .69, .99, .875)
pFA <- c(.33, .45, .22, .34)
data <- cbind(pH, pFA)
data
qnorm(pH) - qnorm(pFA) #should produce the same output as x below

x <- dprime(data, "pH", "pFA")
x # it is the same!

data1 <- cbind(pH1, pFA)
qnorm(pH1) - qnorm(pFA) #oh no, "Inf" is the 4th value because of the 1 in ph1
rm(x)
x <- dprime(data1, "pH1", "pFA")
x # the replacement part of the function is working! x has the same values as before

xtabs(~pH, data=dprime.all)
dprime.all$dprime <- dprime(dprime.all, "pH", "pFA")
summary(dprime.all)
xtabs(~pH, data=data3) # this vector still has 0's and 1's because I didn't create a new vector inside the function, but dprime doesn't have any "Inf" or "Nan", so it worked




```



5) Use boxplots to view distribution of d-prime across contexts and by speaker
```{r}

#get rid of NAs for boxplots
dprime.all.noNA <- dprime.all[!is.na(dprime.all$dprime),]
#reorder context to initial, medial, final
dprime.all.noNA$context <- factor(dprime.all.noNA$context, levels = c("I", "M", "F"))


# dprime by speaker and context
ggplot(dprime.all.noNA, aes(x=context, y=dprime, fill = spkr)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Context") + ylab("D-Prime") + theme_bw()

# dprime by speaker and manner type (pharyngeal, uvular, control) 
ggplot(dprime.all.noNA, aes(x=spkr, y=dprime, fill = pu)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Speaker") + ylab("D-Prime") + theme_bw()

# dprime by speaker, context, and manner type (pharyngeal, uvular, control)
ggplot(dprime.all.noNA, aes(x=spkr, y=dprime, fill = pu)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Speaker") + ylab("D-Prime") + theme_bw() + facet_grid (.~context)

# dprime by speaker and voicing type (voiced, voiceless, control)
ggplot(dprime.all.noNA, aes(x=spkr, y=dprime, fill = vvl)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Speaker") + ylab("D-Prime") + theme_bw()

#dprime by speaker, context, and voicing type (voiced, voiceless, control)
ggplot(dprime.all.noNA, aes(x=spkr, y=dprime, fill = vvl)) + geom_boxplot() + scale_fill_brewer(palette="Set1") + xlab("Speaker") + ylab("D-Prime") + theme_bw() + facet_grid (.~context)

```

